Discrete Quantum Theories and Computing

### 2018-04-10 dissertation.tex
dissertation.lyx "Chapter 1 Introduction" Updating...

\begin{acknowledgments}Thank Prof. Gerardo Ortiz, Prof. Amr Sabry,
and Prof. Andrew Hanson for publishing three papers \cite{geometry2013,DQT2014,THOS2017}
with me, and enormous other things. These papers become the main part
of my thesis, especially most of Chapter~\ref{chap:DQT=000026DQC}
is based on ``Geometry of discrete quantum computing''~\cite{geometry2013}
and ``Discrete quantum theories'' (DQT)~\cite{DQT2014}; most of
Chapter~\ref{chap:DQT=000026DQC} is based on ``Quantum Interval-Valued
Probability: Contextuality and the Born Rule'' (QIVPM) \cite{THOS2017}.
Together with Prof. Dylan Thurston and Prof. Shouhong Wang, thank
you for getting together to understand what I did and passing my dissertation
proposal Spring 2017. Also, thank Prof. Dylan Thurston for serving
in my advisory committee and my Tier 3 committee, with other advisory
committee members passing my Computer Science qualifying exam Spring
2016. Especially, my survey of real computation in the qualifying
exam inspires me the comparison of quantum computer and analog computer
in Chapter~\ref{chap:Introduction}. Chapter~\ref{chap:Introduction}
also contains the motivation in my application for Rethinking Foundations
of Physics 2017 workshop reviewed by Prof. Amr Sabry. Thank Prof.
Lawrence Moss for inviting me to present our results in the interdisciplinary
logic seminar and SICE theory seminar, co-chairing my Tier 3 committee,
writing an assessment letter for me, and drawing my attention on Prof.
Abramsky's paper~\cite{Abramsky2012}, which inspired me to merge
quantum probability with discrete quantum theories. Thank John Gardiner
for inspiring discussion on the difficulties merging them~\cite{Gardiner2014}.
These results are improved and motivated the transition from DQT to
QIVPM in Sec.~\ref{sec:Toward-IVPM}. Thank Prof. Tom Lewis for helping
me organize the unpublished deterministic quantum algorithms in the
term paper for SLST-T501 Academic Writing course as presented in Sec.~\ref{discretequantumtheoryI}.
Thank Traci Nagle for going through the literature review about finite
precision measurement and contextuality discussed for CSCI-Y790 Writing
and Editing course as presented in Sec.~\ref{sec:Quantum-Contextuality}.

@@

analog computing and probabilistic computing

>>>>

\begin{acknowledgments}Thank Prof. Gerardo Ortiz, Prof. Amr Sabry,
and Prof. Andrew Hanson for publishing three papers \cite{geometry2013,DQT2014,THOS2018}
with me, and enormous other things. These papers become the main part
of my thesis, especially most of Chapter~\ref{chap:DQT=000026DQC}
is based on ``Geometry of discrete quantum computing''~\cite{geometry2013}
and ``Discrete quantum theories'' (DQT)~\cite{DQT2014}; most of
Chapter~\ref{chap:DQT=000026DQC} is based on ``Quantum Interval-Valued
Probability: Contextuality and the Born Rule'' (QIVPM) \cite{THOS2018}.
Together with Prof. Dylan Thurston and Prof. Shouhong Wang, thank
you for getting together to understand what I did and passing my dissertation
proposal Spring 2017. Also, thank Prof. Dylan Thurston for serving
in my advisory committee and my Tier 3 committee, with other advisory
committee members passing my Computer Science qualifying exam Spring
2016. Especially, my survey of real computation in the qualifying
exam inspires me the comparison of quantum computer and analog computer
in Chapter~\ref{chap:Introduction}. Chapter~\ref{chap:Introduction}
also contains the motivation in my application for Rethinking Foundations
of Physics 2017 workshop reviewed by Prof. Amr Sabry. Thank Prof.
Lawrence Moss for inviting me to present our results in the interdisciplinary
logic seminar and SICE theory seminar, co-chairing my Tier 3 committee,
writing an assessment letter for me, and drawing my attention on Prof.
Abramsky's paper~\cite{Abramsky2012}, which inspired me to merge
quantum probability with discrete quantum theories. Thank John Gardiner
for inspiring discussion on the difficulties merging them~\cite{Gardiner2014}.
These results are improved and motivated the transition from DQT to
QIVPM in Sec.~\ref{sec:Toward-IVPM}. Thank Prof. Tom Lewis for helping
me organize the unpublished deterministic quantum algorithms in the
term paper for SLST-T501 Academic Writing course as presented in Sec.~\ref{discretequantumtheoryI}.
Thank Traci Nagle for going through the literature review about finite
precision measurement and contextuality discussed for CSCI-Y790 Writing
and Editing course as presented in Sec.~\ref{sec:Quantum-Contextuality}.

@@

When we check the compatibility between quantum mechanics and computer
science, we found their fundamental assumptions are different. On
one hand, the quantum theory is based on infinitely precise real and
complex numbers. Although its predicting ability to quantum phenomenon
is limited by measurement precision, it can well describe the physical
reality together with error analysis. On the other hand, current computers
perform digital computation most of the time, except some analog chips
communicating with physical world~\cite{Camenzind2005}. While the
computability of digital parts is faithfully characterized by its
theoretical model, the Turing machine~\cite{Turing_1937}, the computability
of analog chips in reality is far weaker than the theoretical prediction
of real computation \cite{Siegelmann1998,Ziegler2007,weihrauch2012computable,blum2012complexity}.
One of the reason behind this gap is that our measurement precision
is always finite, and some of the difference cannot be compensated
by error analysis technique. All these problems for classical computation
could potentially apply to quantum computation. To address these issues,
we tried to develop two different types of quantum theories and computing.
Since classical discrete computing model faithfully represents digital
computers, we first tried to build discrete quantum theories and computing
by considering quantum theories and computing over finite fields in
Chapter~\ref{chap:DQT=000026DQC}. In order to , We then tried to
incorporate the idea of finite precision measurement with quantum
theories in Chapter~\ref{chap:QIVPM}.
